Data Analysis and Statistics

### 2025-02-25 Background/DataAnalysisAndStatistics.qmd
EOD push
Adding detail to Data Analysis background page

## TMT quantitative mass spec pipeline

@@

## PSM quantitative mass spec analysis

@@

* Need these details.

>>>>

## TMT quantitative mass spec pipeline {#sec-DataAnalysis_TMTPipeline}

@@

<!-- 
## PSM quantitative mass spec analysis

@@

* Need these details. -->

## Visualizing proteomics data {#sec-DataAnalysis_visualization}

### Volcano Plots

* Visualize enrichment data versus the statistical p value associated with every protein identified in a dataset. 
* In each of the datasets depicted on this site, data points to the right of the x-axis (logFC > 0) denote proteins which were enriched when the lipid probe was irradiated with UV light -- stimulating chemical crosslinking.
* The p value of each protein was -log10 transformed, meaning that the smaller the p value, the higher the point is. 
* In the datasets depicted on this site, there is little meaning to points with logFC < 0; as such, these proteins are ignored regardless of their statistical significance

### Rank-ordered Plots

* Line up each protein identified in the data set from lowest logFC to highest.
* Enables the viewer to see the shape of the dataset -- for example, some datasets will have only a few proteins substantially enriched, and this is reflected by a sharp "elbow" at the edges of the plot. 
* This can be helpful in selecting logFC cutoffs when determining significance cutoffs.

### MA Plots

* Depict the logFC of each protein (**M**) versus the average m/z intensity of the peptides (**A**) that contributed to that proteins identification.
* Gives greater insight into the quality of a datapoint. For example, a point on the far left of the MA plot had relatively low total abundance and may be more sensitive to dramatic changes in enrichment -- whereas points to the right were substantially more abundant in the sample and will be less sensitive to change. 
* Thus, "significantly enriched" proteins with higher **A** intensity may be more trustworthy than those with lower **A** values.

### Heat Maps

### Gene Ontology Overview {#sec-DataAnalysis_GeneOntology}

* Gene Ontology (GO) analyses take a list of proteins and determines whether there is a greater than random enrichment of a biological feature. These features can include pathways, molecular functions, cellular compartments or biological processes. 
* In each of the GO analyses presented on this site, we filtered according to a significance cutoff -- these cutoffs are explained on the respective data pages (either "enriched candidates" or "enriched hits", as in the TMT-based datasets, or unique subsets of probe enrichment, as in the PSM-based datasets). 
* Enriched pathways are given statistical significance by assessing the likelihood that a pathway would be overrepresented through random chance.

#### Gene Ontology Dot Plots

#### Gene Ontology CNET maps
